---
title: "A Comparison of Median and Mean CPUE"
author: "Owain Morgan"
date: "26/10/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Median and Mean CPUE

Catch Per Unit Effort (CPUE)
is casually defined as
the weight of fish caught per hour (effort)
and is expressed in units of kg/hr.

When considering CPUE,
it is common to do so over a year
and so a decision is to be made as to how to take
a representative value of CPUE for trips over the year.

Below is included a time series plot
of two commonly used 'averages', namely
mean (red line) and median (blue line).

```{r, echo=FALSE, warning = FALSE, message=FALSE}
library('tidyverse')
library('lubridate')
cpue_data <- read_csv("_data/sample_landings_data_clean.csv")

#Add a column for Weight/kg
cpue_data <- cpue_data %>%
  mutate(Weight_kg = Weight_g/1000.0)

#Now calculate CPUE for each trip
cpue_data <- cpue_data %>%
  #Group by Year and Trip ID
  group_by(Year,Trip_ID) %>%
  #For each trip, calculate CPUE by dividing weight of catch by time spent
  summarise(Trip_CPUE = sum(Weight_kg) / mean(Effort_Hours))


#I assume we want the time series (to compare mean and median)
#for the year rather than for each individual trip.
cpue_data_ann_median <- cpue_data %>%
  group_by(Year) %>%
  summarise(CPUE_kg_per_hr = median(Trip_CPUE)) 

cpue_data_ann_mean <- cpue_data %>%
  group_by(Year) %>%
  summarise(CPUE_kg_per_hr = mean(Trip_CPUE)) 

cpue_data_ann_mean$Type <- "Mean"

cpue_data_ann_median$Type <- "Median"

cpue <- rbind(cpue_data_ann_mean, cpue_data_ann_median)
#Fortunately cpue_data_ann_median matches with the published results.

ggplot(cpue, aes(x=Year, y=CPUE_kg_per_hr, colour = Type)) +
  geom_point(shape = 24) +
  geom_line() +
  #xlab("Year") +
  #ylab("Comparison of CPUE [kg/hr]") +
  ylab("CPUE [kg/hr]") +
  ggtitle("Comparison of Mean and Median for CPUE") +
  #theme(legend.justification = "left")
  theme(legend.position = c(.05,.95))

```

## A Brief Analysis
As can be seen, 
for years between 2005 and 2012
(where CPUE is relatively small)
the two 'averages' yield
very similar results.

However for years with greater CPUE
such as 2003 and especially for 2013 & 2014,
the mean is significantly higher.
This suggests that in these years, 
there were a greater number of large outliers.
This may have been
due to
a few very efficient trips
which could happen by finding a shoal of 
_Caesio cuning_
and catching a
large number 
thereof.

